### 2021-03-30 comment.tex
blocking arrow, remove paragraphs

\documentclass{article}
\input{definitions}

\begin{document}
\title{An Accurate Model for SARS-CoV-2 pooled RT-PCR Test Errors}

\author[1,2]{Yair Daon}
\author[2,3,*]{Amit Huppert}
\author[1,2,*]{Uri Obolski}

\affil[1]{Porter School of the Environment and Earth Sciences, The
  Faculty of Exact Sciences, Tel Aviv University\\ Tel Aviv, Israel}

\affil[2]{School of Public Health, The Faculty of Medicine, Tel Aviv
  University\\ Tel Aviv, Israel}

\affil[3]{The Gertner Institute for Epidemiology and Health Policy
  Research, Tel Hashomer, Israel}
\affil[*]{Equal contribution}
%\email{yair.daon@gmail.com}
\date{}

\maketitle

\begin{abstract}
\textbf{Background}: Pooling is a popular strategy for increasing
SARS-CoV-2 testing throughput. One popular pooling scheme is Dorfman
pooling: test $N$ individuals simultaneously. If the test is positive
--- retest each individual separately. However, requiring more than
one positive test may lead to increased false-negative rates.

\textbf{Methods}: We analyze the false-negative rate (i.e., the
probability of a negative result for an infected individual) of
Dorfman pooling via a new probabilistic model. We demonstrate that
different, previously made probabilistic assumptions regarding pooling
are unlikely in light of empiric data. Our model is conservative in
that it ignores sample dilution effects, which can only worsen pooling
performance.

\textbf{Results}: We show that one can expect a 60-80\% increase in
false-negative rates under Dorfman pooling, for reasonable parameter
values. Moreover, we show that the false-negative rates under Dorfman
pooling increase when the prevalence of infection decreases.

\textbf{Discussion}: In most pooling schemes, identifying an infected
individual requires positive results in multiple tests and hence
substantially increases false-negative rates. Furthermore, this
phenomenon is more pronounced when infection prevalence is low ---
exactly when pooling is most efficient. Thus, pooling presents an
inherent trade-off: it is most efficient when it is least
accurate. The deterioration of false-negative rates and the
aforementioned trade-off are inherent problems of pooling schemes and
should be kept in mind by practitioners and policy makers.
\end{abstract}
\newpage

\section*{Introduction}
Reverse transcription polymerase chain reaction (RT-PCR) testing is a
key component in breaking transmission chains and mitigating the
COVID-19 pandemic. As such, the need for large-scale testing has
resulted in the development of techniques to increase the the
throughput of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30,
  BayesianDorfman, MatrixPooling, LionDorfman}. These techniques,
often termed \emph{pooling}, or group testing, have originated in the
seminal work of Dorfman and his pooling technique
\cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
selects $N$ individuals and performs a single RT-PCR test on their
combined (\emph{pooled}) samples. If the pooled test yields a positive
result, then each individual is retested separately; otherwise,
everyone is declared negative. The throughput efficiency of Dorfman
pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
its error rates thoroguhly investigated \cite{Kim, Simplistic1,
  OptimalDorfmanPool}.

Studies focused on pooling for SARS-CoV-2 are in a consensus that
sample dilution effects \cite{DilutionHIV, GroupDilution} are not a
concern for pools as large as 64 individuals \cite{PoolSize30, Lion,
  DorfmanYuvalDor, DilutionCOVID}. Consequently, studies assumed that
the probability of a true-positive (the test's \emph{sensitivity})
does not depend on the number of infected samples in the pool, but
rather on the existence of at least one such sample. Thus, the
probability of a positive result in a pooled test has been (assumed)
identical for a pool with one sample from an infected individual and,
e.g., five such samples. This assumption is common in the group
testing literature \cite{Kim, OptimalDorfmanPool}, as well as in more
specific, COVID-19 focused studies \cite{Simplistic1, Simplistic2}. In
this study we challenge these commonly made assumptions and show how
using a more accurate probabilistic model affects estimation of
false-negative rates for Dorfman pooling.

% \subsection{RT-PCR}\label{subsec:rtpcr}
% We briefly explain the process of testing for SARS-CoV-2 by RT-PCR in
% Section~\ref{subsec:rtpcr}. Our exposition intentionally avoids many
% details, and a comprehensive review of RT-PCR can be found in
% \cite{COVID-RTPCR, PCRBook}.

% PCR is a process in which a targeted DNA strand's frequency is
% amplified to create billions of copies in a reaction mixture. At the
% end of this \emph{amplification} process, the presence of targeted DNA
% molecules in a reaction mixture can be confidently
% determined. Arguably, the best way to understand PCR is as a chain
% reaction: Each targeted DNA molecule is copied into two, that are
% copied into four and so forth. This cascade gives PCR its name:
% polymerase chain reaction.

% Specifically, the PCR process is comprised of cycles, with each cycle
% doubling the abundance of the targeted DNA. In each cycle, each
% double-strand DNA molecule is broken into two separate strands. An
% enzyme called DNA polymerase generates a new double-strand DNA replica
% from each separate strand. Replication cannot start without a specific
% short DNA sequence, called a primer, that has to be introduced into
% the reaction mixture. Introducing the right primer into the reaction
% mixture ensures (almost) only the targeted DNA sequence is copied. A
% protein is added to the reaction mixture, which emits light upon
% successfully binding the targeted DNA. Once enough light is emitted
% the tested reaction mixture is declared positive and a
% \emph{detection} is said to occur. If, on the other hand, a predefined
% number of cycles pass without a detection event then the reaction
% mixture is declared negative.

% Since the genetic material of SARS-CoV-2 is RNA rather than DNA, an
% extra preprocessing step is required. An enzyme called reverse
% transcriptase (hence the RT in RT-PCR) replicates existing RNA in the
% reaction mixture into DNA. Once DNA replicas are made, the PCR process
% proceeds as described above, targeting SARS-CoV-2 DNA replicas.

%\subsection{Pooling}
% One possible source of failed detection of SARS-CoV-2 RNA in pooled
% RT-PCR testing (a \emph{false-negative}) is sample dilution. When
% pooling, several samples are mixed, so the concentration of viral RNA
% is reduced. This effect may cause a delay in amplification
% \cite{DorfmanYuvalDor}, no detection and, consequently, a
% false-negative. Another potential source for false-negatives is a
% failed sampling process.

%% However, \cite{Lion} showed that this effect can be
%% safely ignored when mixing up to 32 samples, which we correspondingly
%% assume.

\section*{Methods}
% We define and present the \emph{common assumption} in Section~\ref{subsec:common}
% and refute it in Section~\ref{subsec:refute}. Our model is presented
% in Section~\ref{subsec:ours}.

Formally, we consider a pool containing $N$ individuals
$\{1,\dots,N\}$. We denote the true infection state $\theta \in
\{0,1\}^N$, so individual $i$ is infected iff $\theta_i=1$. The RT-PCR
test's sensitivity (true-negative rate) is denoted $\sens$, and the
test's specificity (true-positive rate) is denoted $\spec$. Pooled
test result (data) is denoted $\data \in \{0,1\}$, where $\data=0$ iff
the test returned a negative result.

\subsection*{The common assumption}\label{subsec:common}
Previous studies of group testing strategies assumed that the
false-negative probability does not depend on the number of infected
samples, but merely on the existence of at least one such sample in a
pool \cite{Kim, OptimalDorfmanPool}. Current studies of pooling in the
context of SARS-CoV-2 also employ similar assumptions
\cite{Simplistic1, Simplistic2}. Specifically, these studies assume
that the probability of a negative result is the same for a pool with
a single sample coming from an infected individual and several
 samples from infected individuals. Explicitly:

\begin{align}\label{eq:common}
  \Pr(\data=0|\theta) = 
  \begin{cases} 
    1-\sens & \exists i\ \text{such that } \theta_i = 1\\
    \spec & \text{otherwise}
  \end{cases} 
\end{align}

Below, we refer to \eqref{eq:common} as the \emph{common
  assumption}. 
%% We refute the assumptions underlying \eqref{eq:common} in
%% Section~\ref{subsec:ours} and suggest more plausible ones in
%% Section~\ref{subsec:ours}.

\subsection*{Refuting the common assumption}\label{subsec:refute}
We refute the common assumption with experimental data collected from
\cite{Salazar}, and summarized in Table \ref{table}. There, the
authors investigate Dorfman pooling and, regardless of the pooled test
result, follow up and test each pool member separately. We focus on
128 pools for which at least one subsequent separate test was positive
--- of which 29 pooled tests were negative and 99 positive. In the
data cited in \cite{Salazar}, of the 29 negative pools, subsequent
separate testing yielded a single positive result in 24. In contrast,
of the 99 positive pools, 42 yielded a single positive test upon
subsequent separate testing.

The data in Table \ref{table} allows us to test the following null
hypothesis $H_0$: The probability of a pooled false-negative is equal
for pools with one subsequent positively tested member and pools with
two or more such members. $H_0$ is a direct consequence of the common
assumption, and rejecting $H_0$ implies the common assumption is not
realistic, at least for SARS-CoV-2.

We apply Fisher's exact test for the presence of more than one
positive individual in correctly identified pools. Fisher's test
yields an increased odds ratio of 6.4, 95\% CI (2.2,23.4), with a
p-value $\approx 10^{-4}$. Thus, we reject $H_0$, refuting the common
assumption.

\begin{table}[h]
\centering
\begin{tabular}{ c c c }
                                & Negative pool  & Positive pool \\%& Row total \\ 
\# subsequent positives $=1$    & 24             & 42            \\%& 66        \\  
\# subsequent positives $\geq1$ & 5              & 57            \\%& 62        \\
 Total                          & 29             & 99            \\%& 128 
\end{tabular}
\caption{Contingency table of data from \cite{Salazar}}\label{table}
\end{table}

\subsection*{Our model}\label{subsec:ours}
Since the essence of the refuted common assumption is that
amplification of all samples occurs only once, we assume the opposite:
amplification of viral RNA succeeds or fails for each sample
independently. According to \cite{Simplistic1, Simplistic2, Kim,
  OptimalDorfmanPool}, a false-positive does not
depend on the number of negative samples in a pool. We incorporate
this assumption in our model with a small modification: We assume that
an erroneous amplification can occur in \emph{any} pool. Specifically,
it is possible that correct amplifications fails and an erroneous one
occurs simultaneously. This assumption is somewhat specific for the
current application of screening for SARS-CoV-2 via RT-PCR. For
example, cross-reactivity with other coronaviruses would have violated
this assumption. However, cross-reactivity was ruled out in
\cite{KitComparison}.
% To summarize, we assume that for a single pool, a positive test result
% is generated in one of two paths. Either, SARS-CoV-2 RNA from an
% infected individual's sample is correctly amplified and detected, and
% this can happen for each positive sample in a pool, with probability
% $\Se$. Or, some erroneous amplification occurs (e.g. contaminant viral
% RNA is introduced), an event that occurs at most once per pool, with
% probability $1-\Sp$.
These assumptions lead to our model, which is illustrated in 
Figure~\ref{fig:likelihood} and summarized in
%% A negative pooled test result occurs when all detection paths (both
%% correct and erroneous) fail. The probability of no false-detection
%% accounts for the $\spec$ term. The probability of no correct detection
%% is $(1-\sens)$ per infected individual. The probability that all such
%% paths fail is the product of the above mentioned terms, displayed in
\eqref{eq:likelihood a}.%%  The probability of a positive result,
%% presented in \eqref{eq:likelihood b}, is simply the complement.
%% The probability of a negative pooled test result is presented in
%% \eqref{eq:likelihood a} (along with its complement
%% \eqref{eq:likelihood b}), and explained below.
%\begin{subequations}
\begin{align}%\label{eq:single likelihood}
  %\begin{split}
    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum_i\theta_i}. \label{eq:likelihood a}
    %
    %
    %
    %% \Pr(\data=1 | \theta) &= 1 - \spec\prod_{i=1}^N
    %% (1-\sens)^{\theta_i} = 1 - \spec(1-\Se)^{\sum\theta_i} . \label{eq:likelihood b}
  %\end{split}
\end{align}
%\end{subequations}
%% Combining \eqref{eq:likelihood a} and \eqref{eq:likelihood b}, and
%% recalling that $\data \in \{0,1\}$ yields:

%% \begin{align}\label{eq:combined likelihood}
%%   \begin{split}
%%     \Pr(\data| \theta) &= \left [\spec\prod_{i=1}^N
%%       (1-\sens)^{\theta_i} \right ]^{1-\data} \left [1 - \spec\prod_{i=1}^N (1-\sens)^{\theta_i} \right ]^{\data}.
%%   \end{split}
%% \end{align}

\begin{figure}[H]
  \centering
  \begin{tikzpicture}
    \begin{scope}[every node/.style={circle,thick,draw}]
      \node[shape=circle,draw=black] (A) at (0,0) {$\theta_1=1$};
      \node[shape=circle,draw=black] (B) at (3,0) {$\theta_2=1$};
      \node[shape=circle,draw=black] (D) at (6,0) {$\theta_3=0$};
      %\node[shape=circle,draw=black] (E) at (6,0) {$\theta_4=1$};

      \node[shape=ellipse,draw=black] (F) at (8,5) {False-positive};

      \node[shape=rectangle,draw=black] (G) at (1.5,5)
           {$\Pr(\data=0|\theta)=(1-\sens)^2\spec$};
    \end{scope}

    \begin{scope}[>={Stealth[black]},
        every node/.style={fill=white},
        every edge/.style={draw=black,very thick}]
      \path [->] (A) edge node {$1-\sens$} (2,4.6);
      %\path [->] (B) edge node {$0$} (G);
      %\path [->] (D) edge node {$0$} (G);
      \path [->] (B) edge node {$1-\sens$} (2.5,4.6);
      \path [->] (F) edge node {$\spec$} (3.7,5);
      \path [-Bar] (D) edge (4.5, 2);
    \end{scope}
  \end{tikzpicture}
  \caption{Illustration of our model. A pool contains individuals
    $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals 1 and 2
    are infected). A negative pooled test ($\data=0$) occurs when
    three detection paths fail: A false-negative occurs for
    individuals $1$ and $2$, each with probability $(1-\sens)$.
    Additionally, no false-positive detection of SARS-CoV-2 occurs,
    with probability $\spec$. Individual $3$ is not infected and does
    not contribute to the probability of the pooled test result. The
    pooled test is positive ($\data=1$) if any path results in a
    detection.}\label{fig:likelihood}
\end{figure}

%% \subsection{Scheme efficiency}
%% The efficiency of Dorfman pooling scheme is the expected number of
%% tests utilized. Let $T$ the number of tests. Then:
%% \begin{align}
%%   \begin{split}
%%     \mathbb{E}[T] &= 1 + N\Pr(\data=1)\\
%%     %
%%     %
%%     %
%%     &= 1 + N\sum_{\theta} \Pr(\theta) \Pr(\data=1|\theta) \\
%%     %
%%     %
%%     %
%%     &= 1 + N\sum_{k=0}^N \Pr(\sum \theta_i = k) \Pr(\data=1|\sum \theta_i = k) \\
%%     %
%%     %
%%     %
%%     &= 1 + N\sum_{k=0}^N \binom{N}{k}q^k(1-q)^k (1-\Sp(1-\Se)^k) \\
%%     %
%%     %
%%     %
%%     &= 1+N - N\Sp\sum_{k=0}^N \binom{N}{k}q^k(1-q)^k (1-\Se)^k \\
%%     %
%%     %
%%     %
%%     %&= 1+N - N\Sp \sum_{k=0}^N \binom{N}{k}(q-q\Se)^k(1-q)^k \\
%%     %
%%     %
%%     %
%%     &= 1+N - N\Sp (1-q\Se)^N\\    
%%   \end{split}
%% \end{align}

%% Compare to the common assumption:

%% \begin{align}
%%   \begin{split}
%%     \mathbb{E}[T] &= 1 + N\Pr(\data=1)\\
%%     %
%%     %
%%     %
%%     &= 1 + N\sum_{\theta} \Pr(\theta) \Pr(\data=1|\theta) \\
%%     %
%%     %
%%     %
%%     &= 1 + N\Big[\Pr(\theta_i = 0,\ \forall i)(1-\Sp) + \Pr(\exists i\ \theta_i =1)\Se\Big] \\
%%     %
%%     %
%%     %
%%     &= 1 + N\Big [(1-q)^N(1-\Sp) + (1-(1-q)^N)\Se \Big ] \\
%%     %
%%     %
%%     %
%%     &= 1 + N\Big [(1-q)^N(1-\Sp - \Se) + \Se \Big ] \\
%%     %
%%     %
%%     %
%%     &= 1 + \Se N + N(1-q)^N(1-\Sp - \Se).
%%   \end{split}
%% \end{align}

%% The difference is:
%% \begin{equation}
%%   \mathbb{E}^{\text{exact}}[T] - \mathbb{E}^{\text{common}}[T] = N \Big [1 - \Sp(1-q\Se)^N - \Se - (1-q)^N(1-\Sp - \Se) \Big ]
%% \end{equation}

\subsection*{Scheme false-negative rate}
We calculate the false-negative rate for a single \emph{infected}
individual, henceforth referred to as "Donald" under our model and
under the common assumption. We distinguish three types of
false-negative events when performing pooling. A \emph{single test}'s
false-negative is the event of a negative result upon testing Donald
separately, i.e., in an RT-PCR test without pooling. A \emph{pooled}
false-negative occurs when a pooled test containing Donald's sample
(and other samples) yields a negative result, i.e., the pooling fails
to detect at least one positive result. Lastly, a \emph{scheme}
false-negative occurs when an entire pooling scheme fails to identify
Donald as infected. Our first task is to calculate Dorfman's scheme
false-negative rate. Equivalently, what is the probability of not identifying Donald as
infected under a Dorfman pooling scheme?

Denote the prevalence of infection in the (tested) population $q$. We
assign Donald as individual $1$, so that $\theta_1=1$. Then:

\begin{align}
  \begin{split}
    \Pr(\data=0|\theta_1=1) &= \sum_{\theta_2,\dots,\theta_N}
    \Pr(\theta_2,\dots,\theta_N)
    \Pr(\data=0|\theta_1=1,\theta_2,\dots,\theta_N) \\
    %
    %
    %
    &= \sum_{k=0}^{N-1}\Pr(\sum_{i=2}^N\theta_i=k)
    \Pr(\data=0|\theta_1=1,\sum_{i=2}^N\theta_i = k)\\
    %
    %
    %
    &= \sum_{k=0}^{N-1}\binom{N-1}{k}q^k(1-q)^{N-1-k} \Sp(1-\Se)^{1+k}\\
    %
    %
    %
    &= \Sp(1-\Se) \sum_{k=0}^{N-1}\binom{N-1}{k}
    \left(q(1-\Se)\right)^k(1-q)^{N-1-k}\\
    %
    %
    %
    &= \Sp(1-\Se)(1-q\Se)^{N-1}. \\
  \end{split}
\end{align}

If the pooled test yields a positive result, Donald is tested
separately. Taking a conservative stand, we assume such a simple
procedure poses no risk of introducing contaminant RNA. Therefore, the
separate test yields a positive result with probability $\Se$.

We calculate the probability that Donald is mistakenly identified as
not infected, henceforth referred to as the scheme's false-negative
rate and denoted $\mi$. In order to correctly identify Donald
as infected, both pooled and separate tests have to yield a positive
result. Thus, the scheme's false-negative rate is:
\begin{align}\label{eq:sfn}
    \begin{split}
        \mi :&= 1 - \Se\Pr(\data=1|\theta=1)\\
        %
        &= 1 - \Se\left [1 - \Sp(1-\Se)(1-q\Se))^{N-1}\right].
    \end{split}
\end{align}

\subsubsection*{Comparison metric}
The single test false-negative rate $1-\Se$ and scheme false-negative
rate $\mi$ are compared via:
\begin{equation}\label{eq:erel}
\err{exact} := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100.
\end{equation}
$\err{exact}$ is the percentage increase in the pooling scheme
false-negative rate, relative to the single test false-negative rate.

According to the common assumption, the scheme false-negative rate is
$1-\Se^2$. A short calculation shows that this approximation implies
the percentage increase in scheme false-negative rate is $\err{common}
:= 100\cdot \Se$.

\section*{Results}\label{section:results}
\subsection*{Scheme false-negative}
We plot $\err{exact}$ for varying prevalence $q$ and sensitivity $\Se$
values, and make the comparison with $\err{common}$. As recommended by
\cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for
different prevalence values. We observe that for a false-positive rate
$\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of reasonable
sensitivity and prevalence values \cite{KitComparison,
  InterpretingCOVID19Test, EstimatingRatesLourenco,
  FalsePositiveEstimate}, an increase of at least $60\%$ in
$\err{exact}$ can be expected (Figure \ref{fig1}).

Interestingly, an increase in infection prevalence monotonically
decreases the scheme false-negative rate, as can also be easily seen
from \eqref{eq:sfn}. For the chosen parameter ranges, the increase in
the single test false-negative rates increases the relative error
$\err{exact}$. These effects can be seen in Figure \ref{fig1} (left
panel), upon conditioning on pool size. Extending the range for $\Sp$
yields no qualitative differences. We further compare $\err{exact}$ to
$\err{common}$, showing the discrepancy changes as a function of both
prevalence and the single test sensitivity (Figure \ref{fig1}, right
panel).
\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{heatmap_sfn.jpg}
  \caption{Relative increase in Dorfman pooling false-negative rates
    $\err{exact}$. Left: Colors represent $\err{exact}$, the relative
    percentage increase in the scheme false-negative rates relative
    to the single test false-negative rates (eq
    \eqref{eq:erel}). Right: colors represent the difference between
    $\err{common}$ and $\err{exact}$. The disease prevalence $q$, is
    varied on the x-axis, while the test sensitivity is varied on
    the y-axis. Pool size $N$, was chosen according to $q$ as in
    \cite{DorfmanYuvalDor}.}\label{fig1}
\end{figure}

\section*{Discussion}
%\subsection{Short recap}
Here, we presented a novel probabilistic model for the outcomes of
pooled PCR tests, parametrized for SARS-CoV-2. Contrary to the common
assumption, we assume, based on data (see section XX), that multiple
infected individuals increase the likelihood of a positive pooled test
result. A direct consequence of our model is that false-negative rates
depend on infection prevalence. Specifically, low values of infection
prevalence increase $\err{exact}$. These results remain qualitatively
similar under varying parameter values, in the observed ranges
\cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
  InterpretingCOVID19Test} (Figure \ref{fig1}).

Our results lead to a conflict: low infection prevalence leads to high
efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, while also
increasing false-negative rates. This is a fact almost disregarded in
\cite{DorfmanYuvalDor}: although (Dorfman) pooling is most efficient
when infection prevalence is low, such circumstances are exactly those
leading to a substantial increase in false-negative rates. Such
behavior of pooling is not implied by the common assumption.

%\subsection{Implications for group testing in different prevalence settings}
As the COVID-19 pandemic progresses, the infection prevalence
in various tested populations undergoes frequent changes. Hence, as
our results suggest, pooling strategies employed for mass testing should be
used with caution in populations where infection rates are low. Such
mass-tested populations often include air travel passengers
\cite{JTM,RobinHood}, potentially importing infections, or
presymptomatic and asymptomatic individuals, and can be crucial for
controlling outbreaks \cite{MinaScience}.

%% One important consequence of the efficiency charts is that
%% practitioners order RT-PCR reagents, they should expect \emph{lower}
%% scheme efficiency, and thus order more reagents.

%\subsection{Limitations}
Lacking large amounts of experimental data, it is hard to assess the
validity of our model. The data in \cite{Salazar} does allow us to
reject the common assumption but not much more. Our model assumes as
little as as possible, given that the standard assumptions are refuted
and dilution effects are not taken into consideration.

%\subsection{Ways to reduce errors}
Although our application focused on Dorfman pooling, our analysis
applies to other pooling schemes. Pooling schemes
(e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
positive pooled results to correctly identify Donald as
infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
If the first pool yields a positive result, it is split in two;
otherwise, everyone are declared negative. Then the splitting is
repeated until resulting pools are negative or individuals are tested
separately. With an initial pool size of 32, Donald will necessarily
have to test positive in pools of size 32, 16, 8, 4 and 2, as well as
in a single test, for the scheme to correctly identify him as
infected. Compare this to the Dorfman scheme that requires a positive
test in a pool of size $N=8$, and an additional single positive test
to identify Donald as infected. The hierarchical pooling scheme of
\cite{Lion, Kim} will necessarily yield more false-negatives than
Dorfman pooling --- there are additional places for it to fail.

Thus, finding improved pooling schemes is imperative. An especially
important consideration in designing such schemes is explicitly taking
the intrinsic PCR error rates into account. We (ref), as well as
others (ref), have developed such models, although their
implementation in real-world settings has not yet commenced.

%\subsection{Conclusion}
We have no doubt that further study of the assumptions underlying
pooling is due. Furthermore, adapting common group testing algorithms
\cite{Kim} to novel error models should be further studied.

%% For example, although Dorfman pooling improves testing throughput,

%% The common assumption implies that one assumes that the probability of
%% a false-negative is identical for the pooled and single
%% tests. Although simplistic in nature, the common assumption does capture
%% the intuition behind our results: For Donald to be considered negative
%% under Dorfman pooling, he has to test negative twice. In addition to
%% this approximation only being an upper bound \cite{Simplistic2}, it
%% does not account for the effect of infection prevalence on the
%% false-negative rates under Dorfman pooling. Supporting the importance
%% of this effect, such an association between prevalence and the scheme
%% false-negative rate under Dorfman pooling has been empirically noted
%% \cite{DorfmanYuvalDor}.

%% Another important factor in pooling is dependence between samples. As
%% mentioned in \cite{DorfmanYuvalDor}, introducing a positive dependence
%% within a pool decreases the false-positive rate. In the extreme case,
%% consider a fully connected pool, where one infection determines the
%% entire pool is infected. In this case, a calculation analogous to the
%% one conducted above (section??) recovers the initial false-negative
%% rate $1-\Se$. Interestingly, pooling was also noted to have increased
%% throughput when infection probabilities are dependent between the
%% pooled individuals \cite{DorfmanYuvalDor}, providing another advantage
%% to sampling dependent individuals in pooling schemes. Hence,
%% deliberately pooling individuals with positive dependence of infection
%% should be further studied and likely encouraged when performing
%% pooling.

%% To conclude, pooling is an important technique that can improve
%% testing throughput in a cost-effective manner. Nevertheless, care must
%% be given to pooling schemes' false-negative rates, especially under
%% low infection prevalence settings.

\bibliographystyle{amsplain}
\bibliography{refs}

\end{document}

%% Todos
%% add refs to dilution and paragraph in start