### 2021-03-31 comment.tex
clean up, remove section on lion

\documentclass{article}
\input{definitions}

\begin{document}
\title{An Accurate Model for SARS-CoV-2 pooled RT-PCR Test Errors}

\author[1,2]{Yair Daon}
\author[1,3,*]{Amit Huppert}
\author[1,2,*]{Uri Obolski}

\affil[1]{School of Public Health, The Faculty of Medicine, Tel Aviv
  University\\ Tel Aviv, Israel}

\affil[2]{Porter School of the Environment and Earth Sciences, The
  Faculty of Exact Sciences, Tel Aviv University\\ Tel Aviv, Israel}

\affil[3]{The Gertner Institute for Epidemiology and Health Policy
  Research, Tel Hashomer, Israel}
\affil[*]{Equal contribution}
\date{}

\maketitle

\begin{abstract}
Pooling is a method of simultaneously testing multiple samples for the
presence of pathogens. Pooling of SARS-CoV-2 tests is increasing in
popularity, due to its high testing throughput. A popular pooling
scheme is Dorfman pooling: test $N$ individuals simultaneously. If the
test is positive, each individual is then tested separately;
otherwise, all are declared negative. Most analyses of pooling schemes
error rates assume that including more than a single infected sample
in a pooled test does not increase the probability of a positive
outcome. We challenge this assumption with experimental data and
suggest a novel and parsimonious probabilistic model for the outcomes
of pooled tests. As an application, we analyze the false-negative rate
(i.e., the probability of a negative result for an infected
individual) of Dorfman pooling. We show that the false-negative rates
under Dorfman pooling increase when the prevalence of infection
decreases. However, low infection prevalence is exactly the condition
when Dorfman pooling achieves highest throughput efficiency. We
therefore implore the cautious use of pooling and development of
pooling strategies that consider the tests' error rates.
\end{abstract}
\newpage

\section*{Introduction}
Reverse transcription polymerase chain reaction (RT-PCR) testing is a
key component in breaking transmission chains and mitigating the
COVID-19 pandemic. As such, the need for large-scale testing has
resulted in the development of techniques to increase the the
throughput of RT-PCR tests \cite{DorfmanYuvalDor, PoolSize30,
  BayesianDorfman, MatrixPooling, LionDorfman}. These techniques,
often termed \emph{pooling}, or group testing, have originated in the
seminal work of Dorfman and his pooling technique
\cite{DorfmanOriginal, DorfmanYuvalDor}. Under Dorfman pooling, one
selects $N$ individuals and performs a single RT-PCR test on their
combined (\emph{pooled}) samples. If the pooled test yields a positive
result, then each individual is retested separately; otherwise,
everyone is declared negative. The throughput efficiency of Dorfman
pooling has been demonstrated empirically \cite{DorfmanYuvalDor} and
its error rates thoroguhly investigated \cite{Kim, Simplistic1,
  OptimalDorfmanPool}.

Studies focused on pooling for SARS-CoV-2 are in a consensus that
sample dilution effects \cite{DilutionHIV, GroupDilution} are not a
concern for pools as large as 64 individuals \cite{PoolSize30, Lion,
  DorfmanYuvalDor, DilutionCOVID}. Consequently, studies assumed that
the probability of a true-positive (the test's \emph{sensitivity})
does not depend on the number of infected samples in the pool, but
rather on the existence of at least one such sample. Thus, the
probability of a positive result in a pooled test has been (assumed)
identical for a pool with one sample from an infected individual and,
e.g., five such samples. This assumption is common in the group
testing literature \cite{Kim, OptimalDorfmanPool}, as well as in more
specific, COVID-19 focused studies \cite{Simplistic1, Simplistic2}. In
this study we challenge these commonly made assumptions and show how
using a more accurate probabilistic model affects estimation of
false-negative rates for Dorfman pooling.

\section*{Methods}
Formally, we consider a pool containing $N$ individuals
$\{1,\dots,N\}$. We denote the true infection state $\theta \in
\{0,1\}^N$, so individual $i$ is infected iff $\theta_i=1$. The RT-PCR
test's sensitivity (true-negative rate) is denoted $\sens$, and the
test's specificity (true-positive rate) is denoted $\spec$. Pooled
test result (data) is denoted $\data \in \{0,1\}$, where $\data=0$ iff
the test returned a negative result.

\subsection*{The common assumption}\label{subsec:common}
Previous studies of group testing strategies assumed that the
false-negative probability does not depend on the number of infected
samples, but merely on the existence of at least one such sample in a
pool \cite{Kim, OptimalDorfmanPool}. Current studies of pooling in the
context of SARS-CoV-2 also employ similar assumptions
\cite{Simplistic1, Simplistic2}. Specifically, these studies assume
that the probability of a negative result is the same for a pool with
a single sample coming from an infected individual and several
 samples from infected individuals. Explicitly:

\begin{align}\label{eq:common}
  \Pr(\data=0|\theta) = 
  \begin{cases} 
    1-\sens & \exists i\ \text{such that } \theta_i = 1\\
    \spec & \text{otherwise}
  \end{cases} 
\end{align}

Below, we refer to \eqref{eq:common} as the \emph{common
  assumption}.

\subsection*{Refuting the common assumption}\label{subsec:refute}
We refute the common assumption with experimental data collected from
\cite{Salazar}, and summarized in Table \ref{table}. There, the
authors investigate Dorfman pooling and, regardless of the pooled test
result, follow up and test each pool member separately. We focus on
128 pools for which at least one subsequent separate test was positive
--- of which 29 pooled tests were negative and 99 positive. In the
data cited in \cite{Salazar}, of the 29 negative pools, subsequent
separate testing yielded a single positive result in 24. In contrast,
of the 99 positive pools, 42 yielded a single positive test upon
subsequent separate testing.

The data in Table \ref{table} allows us to test the following null
hypothesis $H_0$: The probability of a pooled false-negative is equal
for pools with one subsequent positively tested member and pools with
two or more such members. $H_0$ is a direct consequence of the common
assumption, and rejecting $H_0$ implies the common assumption is not
realistic, at least for SARS-CoV-2.

We apply Fisher's exact test for the presence of more than one
positive individual in correctly identified pools. Fisher's test
yields an increased odds ratio of 6.4, 95\% CI (2.2,23.4), with a
p-value $\approx 10^{-4}$. Thus, we reject $H_0$, refuting the common
assumption.

\begin{table}[h]
\centering
\begin{tabular}{ c c c }
                                & Negative pool  & Positive pool \\%& Row total \\ 
\# subsequent positives $=1$    & 24             & 42            \\%& 66        \\  
\# subsequent positives $\geq1$ & 5              & 57            \\%& 62        \\
 Total                          & 29             & 99            \\%& 128 
\end{tabular}
\caption{Contingency table of data from \cite{Salazar}}\label{table}
\end{table}

\subsection*{Our model}\label{subsec:ours}
Since the essence of the refuted common assumption is that
amplification of all samples occurs only once, we assume the opposite:
amplification of viral RNA succeeds or fails for each sample
independently. According to \cite{Simplistic1, Simplistic2, Kim,
  OptimalDorfmanPool}, a false-positive does not
depend on the number of negative samples in a pool. We incorporate
this assumption in our model with a small modification: We assume that
an erroneous amplification can occur in \emph{any} pool. Specifically,
it is possible that correct amplifications fails and an erroneous one
occurs simultaneously. This assumption is somewhat specific for the
current application of screening for SARS-CoV-2 via RT-PCR. For
example, cross-reactivity with other coronaviruses would have violated
this assumption. However, cross-reactivity was ruled out in
\cite{KitComparison}. These assumptions lead to our model, which is illustrated in 
Figure~\ref{fig:likelihood} and summarized in
\eqref{eq:likelihood a}.
\begin{align}
    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N
    (1-\sens)^{\theta_i} = \spec(1-\Se)^{\sum_i\theta_i}. \label{eq:likelihood a}
\end{align}

\begin{figure}[H]
  \centering
  \begin{tikzpicture}
    \begin{scope}[every node/.style={circle,thick,draw}]
      \node[shape=circle,draw=black] (A) at (0,0) {$\theta_1=1$};
      \node[shape=circle,draw=black] (B) at (3,0) {$\theta_2=1$};
      \node[shape=circle,draw=black] (D) at (6,0) {$\theta_3=0$};
      %\node[shape=circle,draw=black] (E) at (6,0) {$\theta_4=1$};

      \node[shape=ellipse,draw=black] (F) at (8,5) {False-positive};

      \node[shape=rectangle,draw=black] (G) at (1.5,5)
           {$\Pr(\data=0|\theta)=(1-\sens)^2\spec$};
    \end{scope}

    \begin{scope}[>={Stealth[black]},
        every node/.style={fill=white},
        every edge/.style={draw=black,very thick}]
      \path [->] (A) edge node {$1-\sens$} (2,4.6);
      %\path [->] (B) edge node {$0$} (G);
      %\path [->] (D) edge node {$0$} (G);
      \path [->] (B) edge node {$1-\sens$} (2.5,4.6);
      \path [->] (F) edge node {$\spec$} (3.7,5);
      %\path [-Bar] (D) edge (4.5, 2);
    \end{scope}
  \end{tikzpicture}
  \caption{Illustration of our model. A pool contains individuals
    $\{1,2,3\}$ with state $\theta=(1,1,0)$ (i.e., individuals 1 and 2
    are infected). A negative pooled test ($\data=0$) occurs when
    three detection paths fail: A false-negative occurs for
    individuals $1$ and $2$, each with probability $(1-\sens)$.
    Additionally, no false-positive detection of SARS-CoV-2 occurs,
    with probability $\spec$. Individual $3$ is not infected and does
    not contribute to the probability of the pooled test result.}\label{fig:likelihood}
\end{figure}

\subsection*{Scheme false-negative rate}
We calculate the false-negative rate for a single \emph{infected}
individual, henceforth referred to as "Donald" under our model and
under the common assumption. We distinguish three types of
false-negative events when performing pooling. A \emph{single test}'s
false-negative is the event of a negative result upon testing Donald
separately, i.e., in an RT-PCR test without pooling. A \emph{pooled}
false-negative occurs when a pooled test containing Donald's sample
(and other samples) yields a negative result, i.e., the pooling fails
to detect at least one positive result. Lastly, a \emph{scheme}
false-negative occurs when an entire pooling scheme fails to
identify Donald as infected. Our first task is to calculate Dorfman's
scheme false-negative rate. Equivalently, what is the probability of
not identifying Donald as infected under Dorfman pooling?

Denote the prevalence of infection in the (tested) population $q$. We
assign Donald as individual $1$, so that $\theta_1=1$. Then:

\begin{align}
  \begin{split}
    \Pr(\data=0|\theta_1=1) &= \sum_{\theta_2,\dots,\theta_N}
    \Pr(\theta_2,\dots,\theta_N)
    \Pr(\data=0|\theta_1=1,\theta_2,\dots,\theta_N) \\
    %
    %
    %
    &= \sum_{k=0}^{N-1}\Pr(\sum_{i=2}^N\theta_i=k)
    \Pr(\data=0|\theta_1=1,\sum_{i=2}^N\theta_i = k)\\
    %
    %
    %
    &= \sum_{k=0}^{N-1}\binom{N-1}{k}q^k(1-q)^{N-1-k} \Sp(1-\Se)^{1+k}\\
    %
    %
    %
    &= \Sp(1-\Se) \sum_{k=0}^{N-1}\binom{N-1}{k}
    \left(q(1-\Se)\right)^k(1-q)^{N-1-k}\\
    %
    %
    %
    &= \Sp(1-\Se)(1-q\Se)^{N-1}. \\
  \end{split}
\end{align}

If the pooled test yields a positive result, Donald is tested
separately. Taking a conservative stand, we assume such a simple
procedure poses no risk of introducing contaminant RNA. Therefore, the
separate test yields a positive result with probability $\Se$.

We calculate the probability that Donald is mistakenly identified as
not infected, henceforth referred to as the scheme's false-negative
rate and denoted $\mi$. In order to correctly identify an infected individual
as infected, both pooled and separate tests have to yield a positive
result. Thus, the scheme's false-negative rate is:
\begin{align}\label{eq:sfn}
    \begin{split}
        \mi :&= 1 - \Se\Pr(\data=1|\theta=1)\\
        %
        &= 1 - \Se\left [1 - \Sp(1-\Se)(1-q\Se))^{N-1}\right].
    \end{split}
\end{align}

\subsubsection*{Comparison metric}
The single test false-negative rate $1-\Se$ and scheme false-negative
rate $\mi$ are compared via:
\begin{equation}\label{eq:erel}
\err{exact} := \frac{\mi - (1-\Se)}{1-\Se} \cdot 100.
\end{equation}
$\err{exact}$ is the percentage increase in the pooling scheme
false-negative rate, relative to the single test false-negative rate.

According to the common assumption, the scheme false-negative rate is
$1-\Se^2$. A short calculation shows that this approximation implies
the percentage increase in scheme false-negative rate is $\err{common}
:= 100\cdot \Se$.

\section*{Results}\label{section:results}
\subsection*{Scheme false-negative}
We plot $\err{exact}$ for varying prevalence $q$ and sensitivity $\Se$
values, and make the comparison with $\err{common}$. As recommended by
\cite{DorfmanYuvalDor}, we apply different pool sizes $N$, for
different prevalence values. We observe that for a false-positive rate
$\Sp=0.95$ \cite{DorfmanYuvalDor} and a range of reasonable
sensitivity and prevalence values \cite{KitComparison,
  InterpretingCOVID19Test, EstimatingRatesLourenco,
  FalsePositiveEstimate}, an increase of at least $60\%$ in
$\err{exact}$ can be expected (Figure \ref{fig1}).

Interestingly, an increase in infection prevalence monotonically
decreases the scheme false-negative rate, as can also be easily seen
from \eqref{eq:sfn}. For the chosen parameter ranges, the increase in
the single test false-negative rates increases the relative error
$\err{exact}$. These effects can be seen in Figure~\ref{fig1} (left
panel), upon conditioning on pool size. Extending the range for $\Sp$
yields no qualitative differences. We further compare $\err{exact}$ to
$\err{common}$, showing the discrepancy changes as a function of both
prevalence and the single test sensitivity (Figure~\ref{fig1}, right
panel).
\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{heatmap_sfn.jpg}
  \caption{Relative increase in Dorfman pooling false-negative rates
    $\err{exact}$. Left: Colors represent $\err{exact}$, the relative
    percentage increase in the scheme false-negative rates relative
    to the single test false-negative rates (eq
    \eqref{eq:erel}). Right: colors represent the difference between
    $\err{common}$ and $\err{exact}$. The disease prevalence $q$, is
    varied on the x-axis, while the test sensitivity is varied on
    the y-axis. Pool size $N$, was chosen according to $q$ as in
    \cite{DorfmanYuvalDor}.}\label{fig1}
\end{figure}

\section*{Discussion}
Here, we presented a novel probabilistic model for the outcomes of
pooled PCR tests, parametrized for SARS-CoV-2. Contrary to the common
assumption, we assume, based on data (Table~\ref{table}), that multiple
infected individuals increase the likelihood of a positive pooled test
result. A direct consequence of our model is that false-negative rates
depend on infection prevalence. Specifically, low values of infection
prevalence increase $\err{exact}$. These results remain qualitatively
similar under varying parameter values, in the observed ranges
\cite{KitComparison,EstimatingRatesKucrika, EstimatingRatesLourenco,
  InterpretingCOVID19Test} (Figure \ref{fig1}).
Our results lead to a conflict: low infection prevalence leads to high
efficiency of Dorfman pooling \cite{DorfmanYuvalDor}, while also
increasing false-negative rates.

As the COVID-19 pandemic progresses, the infection prevalence
in various tested populations undergoes frequent changes. Hence, as
our results suggest, pooling strategies employed for mass testing should be
used with caution in populations where infection rates are low. Such
mass-tested populations often include air travel passengers
\cite{JTM,RobinHood}, potentially importing infections, or
presymptomatic and asymptomatic individuals, and can be crucial for
controlling outbreaks \cite{MinaScience}.

%% Although our application focused on Dorfman pooling, our analysis
%% applies to other pooling schemes. Pooling schemes
%% (e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
%% positive pooled results to correctly identify an infected individual as
%% such. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
%% If the first pool yields a positive result, it is split in two;
%% otherwise, everyone are declared negative. Then the splitting is
%% repeated until resulting pools are negative or individuals are tested
%% separately. With an initial pool size of 32, Donald will necessarily
%% have to test positive in pools of size 32, 16, 8, 4 and 2, as well as
%% in a single test, for the scheme to correctly identify him as
%% infected. Compare this to the Dorfman scheme that requires a positive
%% test in a pool of size $N=8$, and an additional single positive test
%% to identify Donald as infected. The hierarchical pooling scheme of
%% \cite{Lion, Kim} will necessarily yield more false-negatives than
%% Dorfman pooling --- there are additional places for it to fail.

Thus, improving pooling schemes is imperative. An especially important
consideration in designing such schemes is explicitly taking the
intrinsic PCR error rates into account. We \cite{DOPE}, as well as
others \cite{BayesianDorfman}, have developed such models, although
they have not yet been implemented in real-world settings.

To conclude, pooling is an important technique that can improve
testing throughput in a cost-effective manner. Nevertheless, care must
be given to pooling schemes' false-negative rates, especially under
low infection prevalence settings.

\bibliographystyle{amsplain}
\bibliography{refs}

\end{document}

%% Todos
%% add refs to dilution and paragraph in start